---
title: "FIFA World Cup Dashboard"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: cosmo
    logo: favicon_32x32.png
    favicon: favicon_32x32.png
    social: [ "twitter", "linkedin", "menu" ]
---

<style>                     
.navbar {
  background-color:#56042C;
  border-color:white;
}
.navbar-brand {
color:white!important;
}
</style>   

<style type="text/css">

.chart-title {  /* chart_title  */
   font-size: 15px

</style>


```{r global, include = FALSE}

library(flexdashboard)
library(shiny)
library(plotly)
library(reactable)
library(reactablefmtr)
library(worldfootballR)
library(tidyverse)
library(rvest)
library(teamcolors) # NFL team colors and logos
library(extrafont)
library(gt)
library(gtExtras)
library(ggimage)
library(animation)
library(DBI)
library(RSQLite)
library(glue)
library(ggtext)
library(scales)

## Setup steps

# Optional but makes R prefer not to display numbers in scientific notation
options(scipen = 9999)

# Custom ggplot theme (inspired by Owen Phillips at the F5 substack blog)
theme_custom <- function () { 
  theme_minimal(base_size=11, base_family="Chivo") %+replace% 
    theme(
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = 'transparent', color = "transparent")
    )
}


# Define 538 table theme for Reactable table(s) below
theme_538 <- function() {
    reactable::reactableTheme(
        searchInputStyle = list(width = "31%", backgroundColor = "#F9F9F9"),
        style = list(
            fontFamily = "Chivo"
        ),
        headerStyle = list(
            "&:hover[aria-sort]" = list(
                background = "hsl(0, 0%, 80%)"),
            "&[aria-sort='ascending'], &[aria-sort='descending']" = list(
                background = "#555",
                color = "#FFF"
            ),
            borderColor = "#333"
        ),
        borderColor = "#CDCDCD"
    )
}

##### Custom gt table themes for graphics. Inspired by Tom Mock's excellent blog posts #####

gt_theme_538 <- function(data,...) {
    data %>%
        # Add team logos w/ web_image
        text_transform(
            locations = cells_body(
                columns = flag_logo
            ),
            fn = function(x) {
                web_image(
                    url = x,
                    height = 25
                )
            }
        ) %>%
        # Relabel columns
        cols_label(
            flag_logo = ""
        ) %>%
        opt_all_caps()  %>%
        opt_table_font(
            font = list(
                google_font("Chivo"),
                default_fonts()
            )
        ) %>%
        tab_style(
            style = cell_borders(
                sides = "bottom", color = "transparent", weight = px(2)
            ),
            locations = cells_body(
                columns = TRUE,
                # This is a relatively sneaky way of changing the bottom border
                # Regardless of data size
                rows = nrow(data$`_data`)
            )
        )  %>% 
        tab_options(
            column_labels.background.color = "white",
            table.border.top.width = px(3),
            table.border.top.color = "transparent",
            table.border.bottom.color = "transparent",
            table.border.bottom.width = px(3),
            column_labels.border.top.width = px(3),
            column_labels.border.top.color = "transparent",
            column_labels.border.bottom.width = px(3),
            column_labels.border.bottom.color = "black",
            data_row.padding = px(3),
            source_notes.font.size = 12,
            table.font.size = 16,
            heading.align = "left",
            ...
        ) 
}



## Load data

# kaggle dataset for every international match
wc_results <- read_csv("https://raw.githubusercontent.com/martj42/international_results/master/results.csv")


# scrape vegasinsider for WC probabilities
viURL <- "https://www.vegasinsider.com/soccer/odds/world-cup/"

vi_raw <- viURL %>% 
    rvest:: read_html() %>% 
    rvest::html_nodes(".page-main-content li") %>%
    rvest::html_text()

vi_clean <- vi_raw %>% 
    as_tibble() %>% 
    slice(1:2) #only need team win total data from this text

vi_clean <- vi_clean %>% 
    extract(value, 
            into = c("team", "champ_odds"),
            # regex matching for any amount of consecutive non-digits at the start
            regex = "(^\\D+)(.*)", 
            convert = TRUE
    )

vi_clean$team <- str_remove(vi_clean$team, "[+]")
vi_clean$team <- str_remove(vi_clean$team, "[-]")

# create implied odds calculation
vi_clean <- vi_clean %>% 
    type_convert() %>%
  mutate(implied_odds = 1-champ_odds/(champ_odds+100))


# join groups and flags data

groups <- read_csv("Groups.csv")

wc_probs <- left_join(vi_clean, groups, by = c("team" = "Team"))

wc_probs <- wc_probs %>%
    mutate(rank = row_number()) %>%
  mutate('flag_logo' = paste0('https://raw.githubusercontent.com/steodose/World-Cup/main/flags/', team, '.png')) %>%
    mutate(flag_logo = case_when(
    team == "Costa Rica" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/Costa%20Rica.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    team == "Saudi Arabia" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/Saudi%20Arabia.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    team == "USA" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/United%20States.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    team == "South Korea" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/South%20Korea.png",
    TRUE ~ flag_logo
  )) %>%
  relocate(rank, flag_logo, team, group) 


# load matches
matches <- read_csv("https://projects.fivethirtyeight.com/soccer-api/international/2022/wc_matches.csv")

#create unique match ids
matches <- matches |> 
    mutate(unique_match_id = str_c(team1, team2, sep = "-"))

#group matches and compute goals and points
matches_grouped <- matches |> 
    drop_na() |> #drop rows for matches that haven't occurred yet
    group_by(date, unique_match_id, team1, team2) |> 
    summarise(squad1_score = sum(score1, na.rm = TRUE),
              squad2_score = sum(score2, na.rm = TRUE),
              squad1_xg = sum(xg1, na.rm = TRUE),
              squad2_xg = sum(xg2, na.rm = TRUE))

# Process data frame to get one row per team-game
team_games <- matches_grouped %>% 
    # pivot the home and away values to get one row per team
    pivot_longer(contains('team'), names_to = 'home_away', values_to = 'squad', names_prefix = 'team_') %>% 
    # calculate the points awarded (3 for a win, 1 for a draw, 0 for a loss)
    mutate(score = ifelse(home_away == "team1", squad1_score, squad2_score),
           opp_score = ifelse(home_away == "team1", squad2_score, squad1_score),
           xg = ifelse(home_away == "team1", squad1_xg, squad2_xg),
           opp_xg = ifelse(home_away == "team1", squad2_xg, squad1_xg),
           points = case_when(score > opp_score ~ 3,
                              score == opp_score ~ 1,
                              TRUE ~ 0),
           win = ifelse(points == 3, 1, 0))


#calculate team goals and points
team_stats <- team_games |> 
    group_by(squad) |> 
    summarise(gs = sum(score),
              gc = sum(opp_score),
              gd = gs - gc,
              points = sum(points),
              xgf = sum(xg),
              xga = sum(opp_xg),
              xgd = xgf - xga)





```

Inputs {.sidebar}
-----------------------------------------------------------------------

``` {r}

 selectInput("gameInput", label = h3("Select Squad"), 
    choices = unique(team_stats$squad),
    selected="France", multiple = FALSE)

```



Row
-----------------------------------------------------------------------

### Points {.value-box}

```{r}

total_points <- reactive({
  team_stats %>% 
  filter(squad == input$gameInput) %>% 
  summarise(points = sum(points))
})

# Put valueBox() inside renderValueBox({})
renderValueBox({
  valueBox(value = total_points(), 
           caption = "Points",
           color = "#C0C0C0",
           icon = "fa-fire")
})

```


### Goals Scored {.value-box}

```{r}

total_gs <- reactive({
  team_stats2 %>% 
  filter(squad == input$gameInput) %>% 
  summarise(gs = sum(gs))
})

# Put valueBox() inside renderValueBox({})
renderValueBox({
  valueBox(value = total_gs(), 
           caption = "Goals Scored",
           color = "#C0C0C0",
           icon = "fa-futbol")
})

```


### Goals Conceded {.value-box}

```{r}

total_gc <- reactive({
  team_stats %>% 
  filter(squad == input$gameInput) %>% 
  summarise(gc = sum(gc))
})

# Put valueBox() inside renderValueBox({})
renderValueBox({
  valueBox(value = total_gc(), 
           caption = "Goals Conceded",
           color = "#C0C0C0",
           icon = "fa-thumbs-down")
})


```


### Goal Differential{.value-box}

```{r}

total_gd <- reactive({
  team_stats %>% 
  filter(squad == input$gameInput) %>% 
  summarise(gd = sum(gd))
})

# Put valueBox() inside renderValueBox({})
renderValueBox({
  valueBox(value = total_gd(), 
           caption = "Goal Differential",
           color = "#C0C0C0",
           icon = "fa-plus-circle")
})


```


### Expected Goal Differential {.value-box}

```{r}

team_stats2 <- team_stats %>%
    mutate(xgd = round(xgd, digits = 1))

total_xgd <- reactive({
  team_stats2 %>% 
  filter(squad == input$gameInput) %>% 
  summarise(xgd = sum(xgd))
})

# Put valueBox() inside renderValueBox({})
renderValueBox({
  valueBox(value = total_xgd(), 
           caption = "xG Differential",
           color = "#C0C0C0",
           icon = "fa-futbol")
})

```


Row
-----------------------------------------------------------------------

### Latest Vegas Odds {data-width=450}
    
```{r}

# Reactable table setup

wc_probs_table <- reactive({
  wc_probs %>%
        select(rank, flag_logo, team, group, implied_odds)
    })


# render table
renderReactable({
  reactable(wc_probs_table(),
  theme = theme_538,
                  showSortIcon = TRUE,
                  searchable = TRUE,
                  language = reactableLang(
                      searchPlaceholder = "Search for a team"),
                  defaultPageSize = 100,
                  columns = list(
                      rank = colDef(name = "Rank",
                                       align = "center"),
                      team = colDef(name = "Team",
                                    minWidth = 150,
                                       align = "right"),
                      group = colDef(name = "Group",
                                     style = list(borderLeft = "2px solid #555"),
                                     align = "right"),
                      implied_odds = colDef(name = "Implied Odds",
                                            minWidth = 100,
                                            style = color_scales(wc_probs, colors = paletteer::paletteer_d(
                                               palette = "ggsci::amber_material")),
                                            format =  colFormat(percent = TRUE, digits = 1),
                                            align = "right"),
                      
                      ### add logos using embed_img()
                      flag_logo = colDef(
                          name = "",
                          maxWidth = 40,
                          align = "right",
                          cell = embed_img(height = "25", width = "30")
                      )),
  
                  pagination = FALSE,
                  compact = TRUE, 
                  borderless = FALSE, 
                  striped = FALSE,
                  fullWidth = FALSE, 
                  defaultColDef = colDef(align = "center", minWidth = 95)
        ) %>% 
  add_title("2022 FIFA World Cup Odds",
            font_size = 28) %>% 
  add_subtitle("Odds courtesy of Bet365. Click a column header to sort by that statistic.",
               font_size = 14,
               font_weight = "normal") %>% 
  add_source("Data: Bet365/vegasinsider.com")
  
})


```


### Scoring Profiles {data-width=550}
    
```{r}

# make GS vs GC target plot

team_stats2 <- team_stats %>%
  mutate('flag_logo' = paste0('https://raw.githubusercontent.com/steodose/World-Cup/main/flags/', squad, '.png')) %>%
    mutate(flag_logo = case_when(
    squad == "Costa Rica" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/Costa%20Rica.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    squad == "Saudi Arabia" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/Saudi%20Arabia.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    squad == "USA" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/United%20States.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    squad == "South Korea" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/South%20Korea.png",
    TRUE ~ flag_logo
  ))


latest_match <- max(matches_grouped$date) #get latest match date


goals_plot <- team_stats2 %>%
    ggplot(aes(x = gs, y = gc)) +
    geom_image(aes(image = flag_logo), asp = 16/9, position = "jitter") +
    #annotate("text", x = 12, y = 3, label = "Good", color = "red") +
    #annotate("text", x = 12, y = 7, label = "Fun", color = "red") +
    #annotate("text", x = 2, y = 3, label = "Boring", color = "red") +
    #annotate("text", x = 2, y = 9, label = "Bad", color = "red") +
    labs(x = "Goals Scored",
         y = "Goals Conceded",
         caption = "Data: FiveThirtyEight.com\nGraphic: @steodosescu",
         title = "World Cup 2022 Scoring Profiles",
         subtitle = glue("Thru {latest_match} ")) +
    theme_custom() +
    theme(plot.title = element_text(face = "bold")) +
    theme(plot.subtitle = element_markdown()) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
    geom_hline(yintercept = mean(team_stats2$gc, na.rm = T), color = "red", linetype = "dashed", alpha=0.5) +
    geom_vline(xintercept = mean(team_stats2$gs, na.rm = T), color = "red", linetype = "dashed", alpha=0.5) +
    theme(panel.grid.minor = element_blank()) +
    scale_y_reverse()

goals_plot



```



Row {.tabset}
-----------------------------------------------------------------------

### All Matches
    
```{r}


# join in team logos for inclusion in table
all_matches <- matches %>%
    mutate(match = row_number()) %>%
  mutate('flag_logo1' = paste0('https://raw.githubusercontent.com/steodose/World-Cup/main/flags/', team1, '.png'),
         'flag_logo2' = paste0('https://raw.githubusercontent.com/steodose/World-Cup/main/flags/', team2, '.png')) %>%
  select(match, date, flag_logo1, team1, flag_logo2, team2, prob1, prob2, probtie, score1, score2)


#set up table for reactable
all_matches_react <- reactive({
  all_matches %>%
        select(match, date, flag_logo1, team1, flag_logo2, team2, prob1, prob2, probtie, score1, score2)
    })


# render table
renderReactable({
  reactable(all_matches_react(),
  theme = theme_538,
  columnGroups = list(
                      colGroup(name = "Via FiveThirtyEight.com", 
                               columns = c("prob1", "prob2", "probtie"))
                  ),
                  showSortIcon = TRUE,
                  searchable = TRUE,
                  language = reactableLang(
                      searchPlaceholder = "Search for a team"),
                  defaultPageSize = 100,
                  columns = list(
                      match = colDef(name = "Match",
                                       align = "center"),
                      date = colDef(name = "Date",
                                    minWidth = 120,
                                       align = "center"),
                      team1 = colDef(name = "Home",
                                    minWidth = 120,
                                       align = "right"),
                      team2 = colDef(name = "Away",
                                    minWidth = 120,
                                       align = "right"),
                      prob1 = colDef(name = "Home Probs",
                                            minWidth = 100,
                                            style = color_scales(all_matches, colors = paletteer::paletteer_d(
                                               palette = "ggsci::amber_material")),
                                            format =  colFormat(percent = TRUE, digits = 1),
                                            align = "right"),
                       prob2 = colDef(name = "Away Probs",
                                            minWidth = 100,
                                            style = color_scales(all_matches, colors = paletteer::paletteer_d(
                                               palette = "ggsci::amber_material")),
                                            format =  colFormat(percent = TRUE, digits = 1),
                                            align = "right"),
                       probtie = colDef(name = "Tie Probs",
                                            minWidth = 100,
                                            style = color_scales(all_matches, colors = paletteer::paletteer_d(
                                               palette = "ggsci::amber_material")),
                                            format =  colFormat(percent = TRUE, digits = 1),
                                            align = "right"),
                      score1 = colDef(name = "Home Score",
                                    minWidth = 80,
                                       align = "right"),
                      score2 = colDef(name = "Away Score",
                                    minWidth = 80),
                      
                      ### add logos using embed_img()
                      flag_logo1 = colDef(
                          name = "",
                          maxWidth = 40,
                          align = "right",
                          cell = embed_img(height = "25", width = "30")
                      ),
                      flag_logo2 = colDef(
                          name = "",
                          maxWidth = 40,
                          align = "right",
                          cell = embed_img(height = "25", width = "30")
                      )),
  
                  pagination = FALSE,
                  compact = TRUE, 
                  borderless = FALSE, 
                  striped = FALSE,
                  fullWidth = FALSE, 
                  defaultColDef = colDef(align = "center", minWidth = 95)
        ) %>% 
  add_title("2022 FIFA World Cup Matches",
            font_size = 28) %>% 
  add_subtitle("Odds courtesy of FiveThirtyEight. Click a column header to sort by that statistic.",
               font_size = 14,
               font_weight = "normal") %>% 
  add_source("Data: Bet365/vegasinsider.com")
  
})


```


### Group Stage

```{r}

matches_groups <- matches %>%
    filter(date <= '2022-12-02')

matches_groups_grouped <- matches_groups |> 
    drop_na() |> #drop rows for matches that haven't occurred yet
    group_by(date, team1, team2) |> 
    summarise(squad1_score = sum(score1, na.rm = TRUE),
              squad2_score = sum(score2, na.rm = TRUE),
              squad1_xg = sum(xg1, na.rm = TRUE),
              squad2_xg = sum(xg2, na.rm = TRUE))

team_games_groups <- matches_groups_grouped %>% 
    # pivot the home and away values to get one row per team
    pivot_longer(contains('team'), names_to = 'home_away', values_to = 'squad', names_prefix = 'team_') %>% 
    # calculate the points awarded (3 for a win, 1 for a draw, 0 for a loss)
    mutate(score = ifelse(home_away == "team1", squad1_score, squad2_score),
           opp_score = ifelse(home_away == "team1", squad2_score, squad1_score),
           xg = ifelse(home_away == "team1", squad1_xg, squad2_xg),
           opp_xg = ifelse(home_away == "team1", squad2_xg, squad1_xg),
           points = case_when(score > opp_score ~ 3,
                              score == opp_score ~ 1,
                              TRUE ~ 0),
           win = ifelse(points == 3, 1, 0),
           draw = ifelse(points == 1, 1, 0),
           loss = ifelse(points == 0, 1, 0))


team_stats_groups <- team_games_groups |> 
    group_by(squad) |> 
    summarise(gs = sum(score),
              gc = sum(opp_score),
              gd = gs - gc,
              points = sum(points),
              xgf = sum(xg),
              xga = sum(opp_xg),
              xgd = xgf - xga,
              wins = sum(win),
              draws = sum(draw),
              losses = sum(loss))


joined_df <- left_join(team_stats_groups, groups, by = c("squad" = "Team"))

joined_df <- joined_df %>%
    mutate(rank = row_number()) %>%
  mutate('flag_logo' = paste0('https://raw.githubusercontent.com/steodose/World-Cup/main/flags/', squad, '.png')) %>%
    mutate(flag_logo = case_when(
    squad == "Costa Rica" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/Costa%20Rica.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    squad == "Saudi Arabia" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/Saudi%20Arabia.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    squad == "USA" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/United%20States.png",
    TRUE ~ flag_logo
  )) %>%
    mutate(flag_logo = case_when(
    squad == "South Korea" ~ "https://raw.githubusercontent.com/steodose/World-Cup/main/flags/South%20Korea.png",
    TRUE ~ flag_logo
  )) %>%
  relocate(rank, flag_logo, squad, group) %>%
    arrange(group, desc(points), desc(gd))

# filter for proper groups
groups_ad <- joined_df %>%
    filter(group == 'A' | group == 'B' | group == 'C' | group == 'D') %>%
    #arrange(desc(points)) %>%
    #mutate(rank = row_number()) %>%
    select(flag_logo, squad, group, points, wins, losses, draws, gs, gc, gd, xgd) %>%
    gt(groupname_col = "group") %>%
    gt_theme_538() %>%
    #gt_img_rows(columns = flag_logo, height = 25) %>%
  cols_label(
    flag_logo = "",
  ) %>% 
  cols_align(align = "center",
               columns = 2:6) %>%
  tab_header(title = md("**2022 World Cup Group Stage**"),
               subtitle = glue("Groups A thru D (Nov 18 - Dec 02)"))

#group_ad

groups_eh <- joined_df %>%
    filter(group == 'E' | group == 'F' | group == 'G' | group == 'H') %>%
    #arrange(desc(points)) %>%
    #mutate(rank = row_number()) %>%
    select(flag_logo, squad, group, points, wins, losses, draws, gs, gc, gd, xgd) %>%
    gt(groupname_col = "group") %>%
    gt_theme_538() %>%
    #gt_img_rows(columns = flag_logo, height = 25) %>%
  cols_label(
    flag_logo = "",
  ) %>% 
  cols_align(align = "center",
               columns = 2:6) %>%
  tab_header(title = md("**2022 World Cup Group Stage**"),
               subtitle = glue("Groups E thru H (Nov 18 - Dec 02)"))


tables_combined <- list(groups_ad, groups_eh)

gt_two_column_layout(tables_combined)
    
    
    


```


```{css}
    .chart-shim {
      overflow: auto;
    }
    ```

### About

**World Cup Dashboard**

This dashboard looks at World Cup win probabilities for each match and squad, and overall team stats. App created by [Stephan Teodosescu](https://stephanteodosescu.com/personal-website/).


